Associative Data Structures for Website Filtering

### 2014-04-17 paper/paper.tex
Fixed up results.

    We performed tests using subsets of twelve different sizes, all of the form $2^x$ and with $8 <
    x \le 20$.  The word list was obtained by taking the system's dictionary file and selecting all
    words that contained only lowercase letters and that were at least 8 and at most 12 letters
    long.  The top-level domain list was obtained from Wikipedia, and filtered down to only those
    domains that consisted of two letters from the English alphabet.  The resulting domains were of
    the form \texttt{www.\$word.\$tld}, where \texttt{\$word} is one of words from the the word list
    and \texttt{\$tld} is a top level domain.  For the full paths, a number of domains were selected
    and each was appended a random alphanumeric string of eighty characters.  By using the same
    domain with multiple random strings, a sufficiently large number of paths was generated.
    Finally, as described in the research plan, consecutive 32-bit integers starting from zero were used
    for the IP addresses.

@@

    Our results were great!  They are best analysed with a two-sample independent T test.  We had an
    equal number of measurements about group 1 (with mean $x_1$ and standard deviation $\sigma_1$ and
    about group 2 (with mean $x_2$ and standard deviation $\sigma_2$).  From this we can conclude that
    data structure 1 is, on average, unquestionably superior.

@@

    In particular, constructing a small hash table\footnote{state the number of elements in a
    "small" hash table} after constructing many big ones takes a disproportionally long time.  Only
    the first hash table is affected, and the time is comparable to the time required to construct a
    hash table with 4000 elements, which is illogical.  We assume this is caused by the allocator
    but the reason for such behaviour is still unclear.  The fact that a hashtable needs contiguous
    memory is likely to have some influence, as well as the allocator returning memory to the
    operating system.  However, we reason the latter should also have happened with the other
    structures.

    A similar issue occurs around 4000 or 8000 elements (depending on data structure).  For all
    three datastructures, the cost of inserting and querying grows at a higher rate than before.
    This is difficult to quantify\footnote{Can we do some sort of linear regression to prove this is
    the case?}, but it appears some extra load is added at that point.  We suspect it is related
    either to the allocator having to acquire more memory or the CPU cache running out and more
    cache misses occuring.  We would measure this as well, if there was more time.

    A potentially related issue is that the standard deviation of our measurements rises around the
    same point.  That increases the likelihood of it being allocator-related, as that involves a
    system call and could cause scheduling issues\footnote{elaborate/ref?}, despite the system being
    otherwise very close to idle.

    Also, we acquired counterintuitive results showing that inserting or querying a full path is
    faster than querying only the domain name.  This can partially be explained\footnote{Not really,
    but ehhh.}by the fact that the string implementation we used was based on reference counting.
    %(If it'd also provide SSO, this would be easy to explain, but the darn thing doesn't.)

@@

>>>>

    We performed tests using subsets of thirteen different sizes, all of the form $2^x$ and with $8
    \le x \le 20$.  The word list was obtained by taking the system's dictionary file and selecting
    all words that contained only lowercase letters and that were at least 8 and at most 12 letters
    long.  The top-level domain list was obtained from Wikipedia, and filtered down to only those
    domains that consisted of two letters from the English alphabet.  The resulting domains were of
    the form \texttt{www.\$word.\$tld}, where \texttt{\$word} is one of words from the the word list
    and \texttt{\$tld} is a top level domain.  For the full paths, a number of domains were selected
    and each was appended a random alphanumeric string of eighty characters.  By using the same
    domain with multiple random strings, a sufficiently large number of paths was generated.
    Finally, as described in the research plan, consecutive 32-bit integers starting from zero were
    used for the IP addresses.

@@

    Having done the tests, we have obtained a lot of data that we can base our conclusions on.  A
    preliminary overview can be obtained in Appendix~\ref{App:AveResults}, which lists the average
    per-element results we obtained.  However, in order to draw conclusions from our data we will
    require statistical methods.  For our results, we shall make use of the independent samples
    T-test and the bivariate correlation test.  In all cases, we will use a significance level of
    $0.05$ and perform the tests using the SPSS software package.

    Note that in the following analyses, we do not consider the results of the $2^8$ group, as the
    experimental errors involved were too significant.

    \subsection{Correlation between number of entries and results}

    Firstly, we shall look at the influence the number of entries has on the per element speed of
    operations.  For this we will use a bivariate correlation test, since we want to check if there
    exists a correlation between two variables of ratio levels of measurements.

    Using the analysis functions of SPSS shows that for almost all situations there exists a
    positive correlation between the number of entries and the duration of both queries and inserts.
    For both the skip list and the binary search tree, these correlations are significant for all
    key types. In all cases they have positive correlation coefficients between $0.4$ and $0.6$.

    In the case of hash table, however, we do not always find a correlation.  For all situations there does
    exist a positive correlation between the number of entries and the per element query duration, but
    the insertion duration is only correlated to the number of entries in the case of the IP address
    key type.  For the domain and full path key types, no correlation between the number of entries
    and the duration of insertion operations is observable.

    \subsection{Difference in performance between different keytypes}

    In order to answer the second part of our research question, we must compare the behaviour of
    the data structures given different key types.  We start by fixing the number of entries; we
    chose to do this at 4096 elements.  We then again performed an independent samples T-test
    using SPSS.

    The results are largely the same for all three data structures. For all of the data structures
    the use of IP addresses gives significantly higher speeds both for query and insertion, and the
    memory usage is lower.  These results can be replicated for all entry counts.

    The difference between the results for the full path and the domain key types is less clear-cut.
    In the case of the binary search tree, the insertion speed for full paths is significantly higher
    than the insertion speed for domains, while the opposite is true for query speeds.  The skip
    list also has a significantly higher insertion speed for full paths than for domains, but there
    is no significant difference for query speed unless we increase the number of elements.

    The hash table displays the expected behaviour, and both the insertion and query speeds for
    domains are significantly higher than for full paths.

    \subsection{Comparison between different data structures}

    We would like to be able to use our results to determine which data structures are better in
    which cases.  For this we will have to compare the different data structures for specific
    situations. We will use an independent samples T test again for this, since our data points are
    independent of each other.

    For all element counts and domain or full path key types, the hash table has significantly higher insertion
    and query speeds than the binary search tree and skip list.  For IP addresses, the results are
    more interesting.  While the hash table is still significantly faster for query speeds for any
    number of entries, the binary search tree wins on insertion speed when there are 4096 entries or
    less.

@@

    In particular, constructing a small hash table after constructing many big ones reliably took a
    disproportionally long time.  Only the first hash table is affected, and the time is comparable
    to the time required to construct a hash table with 4000 elements.  We assume this is caused by
    the allocator but the reason for such behaviour is still unclear.  The fact that a hash table
    needs contiguous memory is likely to have some influence, as well as the allocator returning
    memory to the operating system.  However, we reason the latter should also have happened with
    the other structures.

    A similar issue occurs around 4000 or 8000 elements (depending on data structure).  For all
    three datastructures, the cost of inserting and querying grows at a higher rate than before.
    This is difficult to quantify, but it appears some extra load is added at that point.  We suspect it is related
    either to the allocator having to acquire more memory or the CPU cache running out and more
    cache misses occuring.

    This is also a possible explanation that the results we obtained which claimed that operations
    for a full path key are more efficient than identical operations for a domain key.  An
    alternative explanation is that the differences in the data sets used for the keys made
    the comparison cheaper for the full paths, despite their greater size.  However, the way we
    constructed the sets would suggest the opposite.

@@

    \newpage
    \appendix
    \section{Average Results} \label{App:AveResults}

    \begin{figure}[h!]
    \centering
    \includegraphics[width=\textwidth]{ip-address-averages.png}
    \caption{Per element averages with the IP-address key type.}
    \end{figure}

    \begin{figure}[h!]
    \centering
    \includegraphics[width=\textwidth]{domain-averages.png}
    \caption{Per element averages with the domain key type.}
    \end{figure}

    \begin{figure}[h!]
    \centering
    \includegraphics[width=\textwidth]{full-path-averages.png}
    \caption{Per element averages with the full path key type.}
    \end{figure}
